---
title: "assignment RepDate"
author: "koen klein willink"
date: "26 mei 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Step 1: Retrieve data from the web
The first step is the retrieve the right data set from the web. The script is added below. 

```{r retrieve data}
temp <- tempfile()
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip", 
              temp)
con <- unz(temp, "activity.csv")
dat <- read.table(con, header=T, sep=",")
unlink(temp)
```

## Step 2: process/transform data
Ensure the data fields are in the date formate

```{r process/transform}
library(lubridate)
dat$date<-as_date(dat$date)
head(dat)
```

## Step 3: 
Calculate the total steps per day and make plot
```{r total steps per day}
library(dplyr)
stepspd<-dat %>%
  #per date herewith drop the interval
  group_by(date) %>%
  #summarize total steps by taking the sum of steps per    #date
  summarize(totalstep=sum(steps))

#make plot
library(ggplot2)
#define ggplot
plot<-ggplot(stepspd,aes(x=date,y=totalstep))
#characteristics plot choose identity since already summarize date (otherwise a hist is plotted)
plot+geom_bar(stat="identity",color="black")+geom_hline(yintercept=mean(stepspd$totalstep,na.rm=TRUE))
```

##Step 4: Calculate mean and median per day

```{r mean and median per day}
meanpd<-dat%>%
  group_by(date)%>%
  summarize(mean=mean(steps))%>%
  head()

medianpd<-dat%>%
  group_by(date)%>%
  summarize(median=median(steps))%>%
  head()
```

## Step 5: Timeserie plot
Plot timeseries steps per 5 minutes
```{r timeseries}
ts<-dat%>%
  group_by(interval)%>%
  summarize(average=mean(steps,na.rm=TRUE))
tsplot<-ggplot(ts,aes(x=interval,y=average))
tsplot+geom_line()
```

## Step 6: max number of steps per interval
```{r max number}
maximuminterval<-which.max(ts$average)
absolutemax<-max(ts$average)
cat("interval with maximum value is",maximuminterval)
cat(", maximum value is", absolutemax)
```

## Step 7:imputing missing values
first calculate the number of mission values and the percentage of missing value
replacing NA with mean for the interval
via merging and than mutate functions

```{r imputing mission values}
numberofmissingvalues<-count(dat[!complete.cases(dat),])
percentageofmissingvalue<-numberofmissingvalues/count(dat)
cat("number of missing values are", as.character(numberofmissingvalues))

cat(" ,percentage of missing values is", as.character(percentageofmissingvalue*100),"%")

#create identical dataframe with NA replace
new7<-dat%>%
  group_by(interval)%>%
  summarize(mean=mean(steps,na.rm=TRUE))%>%
  merge(dat)%>%
  mutate(steps=ifelse(is.na(steps),mean,steps))%>%
  group_by(date)%>%
  summarize(totalstep=sum(steps))

#make plot
library(ggplot2)
#define ggplot
plot5<-ggplot(new7,aes(x=date,y=totalstep))
#characteristics plot choose identity since already summarize date (otherwise a hist is plotted)
plot5+geom_bar(stat="identity",color="black")+geom_hline(yintercept=mean(new7$totalstep))



```

##Step 8:Plot the difference for weekend and workdays
create identifier for weekends via wday() function and ifelse

```{r workday weekenddays}
new6<-dat%>%
  group_by(interval)%>%
  summarize(mean=mean(steps,na.rm=TRUE))%>%
  merge(dat)%>%
  mutate(steps=ifelse(is.na(steps),mean,steps))%>%
  mutate(wd=wday(date))%>%
  mutate(ident=ifelse(wd==1|wd==7,"weekend","weekday"))%>%
  group_by(interval,ident)%>%
  summarize(mean=mean(steps))

#plot
plot2<-ggplot(new6,aes(x=interval,y=mean))
plot2+geom_line()+facet_grid(ident~.)

```

  